---
title: "Innovation AI - Week1 - EDA"
author: "Yilin Wang"
date: "2026-02-04"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
rm(list = ls())  # Clear workspace

library(tidyverse)
library(readxl)
library(janitor)
library(skimr)
```

Data Load: set the data path and view the names of all sheets from the EXCEL table.
```{r load data, message=FALSE, warning=FALSE}
data_path = "../data/ai_interview_dataset_1129.xlsx"

sheets <- excel_sheets(data_path)
print("Sheets in the dataset:")
sheets           # List all sheet names
```

Read each sheet from the excel file:
```{r read data, message=FALSE, warning=FALSE}
data_list <- list()
for (sheet in sheets) {
  data_list[[sheet]] <- read_excel(data_path, sheet = sheet)
}
```

Dataset Overview: summarises the basic structure of each dataset, including the number of rows, columns, variable names, and missing rate.
```{r overview data, results='asis'}
# Overview of each sheet
overview <- list()
for (sheet in sheets) {
    df <- data_list[[sheet]]
    overview[[sheet]] <- tibble(
        rows = nrow(df),
        cols = ncol(df),
        col_names = paste(names(df), collapse = ", "),
        missing_rate = sum(is.na(df)) / (nrow(df) * ncol(df))
)
}
print("Sheet overview:")
overview
```
From the overview, we can see that the general size, variable name and completeness of each table.


Numeric Variable Summary: select only numeric variables and analyse numeric variables using descriptive statistics.
```{r setup, message=FALSE, warning=FALSE}
summary_list <- list()
for (sheet in sheets) {
    df <- data_list[[sheet]]

    # Select only numeric columns for summary
    numeric_df <- df %>% select(where(is.numeric)) 
    # Generates information such as mean, median, standard deviation, minimum, maximum, and missing values.
    summary_list[[sheet]] <- skim(numeric_df)
}
summary_list
```
The results include mean, median, standard deviation, minimum, maximum, and missing value information for each numeric variable.



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
